Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-08-15 doc/manuscript/manuscript.Rmd
Address comments up to Discussion

<!-- SEE: Avoid sentences that start with the authors/reference. -->
@attanayake2013sclerotinia studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA.
There were no shared microsatellite multilocus genotypes (MLG) between the two populations.
Sexual outcrossing was detected in both populations with higher recombination rate in Chinese population than the US population.
Additionally, phenotypic traits such as production of sclerotia, mycelial compatibility group (MCG), growth rate were also significantly different between the two populations.
Genetic diversity of *S. sclerotiorum* from soy bean, canola, sunflower, and dry bean in the North Central region of the USA was studied by [@aldrich-wolfe2015genetic].
Microsatellite analysis revealed little host- specificity or genetic-isolation of *S. sclerotiorum* isolates among host crops as 93% of the observed genetic variation was explained by within-host differences.
Isolates from all host crops were primarily reproducing clonally while there were also evidence of genetic recombination of isolates from canola and sunflower.

<!-- Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe. 
In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
No white mold resistant dry bean cultivars exist.
Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic]. -->
To date, no commercial white mold-resistant dry bean cultivars exist. 
Field trials in some regions of the US have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation; @steadman2004identification].
<!-- This puny paragraph needs a lot of work to more fully describe the operation of the wmn -->

@@

@@

## Data processing {-}

All analyses were performed on this data set unless otherwise mentioned.
All data processing and analyses were performed in R version 3.4.1 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif. 
<!-- add something about repeat length consistency  -->
Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of region/source/host/year---meaning that one representative of each genotype was retained for each combined regions, source, host, and year---for subsequent analysis.

We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to assess differentiation associated with wmn by quantifying the differentiation conferred by source and region [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
Source in this case was defined as a population belonging to a wmn or not and region defined as state.

To identify specific regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet]. 
This method decomposes the genetic data into principal components, and then uses a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant]. 
To avoid over-fitting the data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
This function implements a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximizes the rate of successful group reassignment. 
We additionally assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis]. 
To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
Both genotypic diversity and linkage disequilibrium were calculated in *poppr* version 2.4.1 [@kamvar2014poppr].

### Cultivar Differentiation {-}

To assess if cultivar had an influence on genetic diversity across wmn, we first subset the clone-corrected data to contain only samples from wmn and from the cultivars Beryl, Bunsi, and G122 and tested differentiation using AMOVA and DAPC as described above.

### Shared Multilocus Genotype Community Analysis {-}

We wanted to evaluate patterns of connectivity between shared multilocus genotypes across geographic regions.
We first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
Each node was weighted by the fraction of shared MLGs.
Each edge represented a single MLG, but because a single MLG could cross populations, that MLG could be represented by (n\*(n-1))/2 edges where n represented the number of populations it crossed.
Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
The graph was plotted using the R packages *ggplot2* version 2.2.1 [@ggplot2] and *ggraph* version 1.0.0 [@ggraph].
To ensure that we captured the same community signal, we additionally performed this analysis including the 5 polymorphic markers described above.

A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in 11 states in the US, Australia, Mexico, and France (Table \@ref(tab:isolate-table)).
Over 50% of the isolates came from four states, MI (62), ND (60), WA (59), NE (47). 
Four regions had fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1).
We observed 87 mycelial compatibility groups (MCG), the most abundant of which (MCG 5) represented 73 isolates over 37 multilocus genotypes (MLG) (Fig \@ref(biggraph)A).
After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with 318 isolates representing 165 MLGs (Table \@ref(tab:iatab)).
Analysis of the genotype accumulation curve showed no clear plateau, indicating that we would likely obtain more multilocus genotypes if we were to collect more loci. The number of observed alleles per locus ranged from 2 to 10 with an average of 6.27 (Table \@ref(tab:locus-stats)).

@@

The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
The forward-backward selection process revealed year, region, host, and aggressiveness to be the optimal variables for the reduced model, accounting for 45% of the total variation.
ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.0674 (P = 0.001).
Assessment of the marginal effects showed that region, host, year, and aggressiveness significantly explained genetic variation (P = 0.001, P = 0.009 for region).
It was revealed that there was multicollinearity combining all variables with MCG, so we followed the same procedure, dropping MCG from the list of potential predictors. 
From these results, we found that aggressiveness, year, region, and host were optimal, accounting for 17.6% of the total variation.
ANOVA revealed significant effects with an adjusted R^2^ of 0.0325 (P = 0.001).
While the marginal effect assessment showed that region, host, and year all significantly explained the variation at P = 0.001, and aggressiveness significantly explained the variation at P = 0.045.
Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
Variance partitioning of the independent variables without MCG indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.

@@

![
Biplot showing five most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
Points represent the 318 clone-corrected samples in ordination space. 
](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=61%}

@@

![
Boxplot representing aggressiveness (vertical axis) by population (horizontal axis) arranged in descending order of mean aggressiveness for all populations with N > 5. Box boundaries represent 25% and 75% percentiles and whiskers extend to at most 1.5 $\times$ the interquartile range. Horizontal line represents median value. Points represent individual observations whose colors indicate if they came from white mold screening nurseries (wmn) or not (other).
](../../results/figures/publication/aggressiveness.pdf){#aggressiveness-boxplot}

@@

Our assessment of aggressiveness in association with Region showed a significant effect (P < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig. \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)). 
MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 (MCG 44) to 4.6 (MCG 49).
We additionally found a significant effect for MLGs (P = 7.44e^-4^) for MLGs, with means that ranged from 6.0 (MLG 78) to 4.3 (MLG 140).

@@

Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph), \@ref(fullgraph)). 
Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set.
The three communities with the most members contained 8 of the ten most abundant MCGs. 
Comparing these communities with Bruvo's genetic distance revealed and average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
When we assessed the number of potential heterothallic parings between different MLGs in our data set (two different MLGs that are in the same mycelial compatibility group), we found an average of 16.5 potential heterothallic parings per MLG. 
MLG 75, representing just four isolates, had 57 neighbors that shared the same mycelial compatibility group. 
Overall, there was no clear pattern to the association between MLG and MCGs.

@@

![
Data show complex associations between Mycelial Compatibility Groups and Multilocus Genotypes.
**A)** Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
Singletons (46) have been removed for clarity, leaving 41 MCGs.
White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG.
**B)** Full graph-representation of the relationship between MCGs (open circles) and MLGs (filled circles). 
Details in fig. \@ref(fullgraph). 
**C)** A subset of **B** representing the 8 most common MCGs and their associated MLGs.
Node size reflect the number of samples represented by each node (circle).
Edges (arrows) point from Multilocus Genotype to Mycelial Compatibility Group where the weight (thickness) of the edge represents the number of shared isolates.
Edges extending from MLGs displayed to other MCGs are not shown.
](../../results/figures/publication/mcg-subgraph-with-context.pdf){#biggraph width=100%}

@@

@@

We observed a total of 165 multilocus genotypes (MLG) (215 with 16 loci).
The most abundant MLG represented 27 isolates (Table \@ref(tab:mlg-table)) across five regions (NE, MI, WA, CO, and ND). 
The genotypes were relatively evenly distributed within regions with moderate to high diversity (Table \@ref(tab:iatab)). 
Of the 165 MLGs, 76 (46%) were found across all geographic regions except WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig. \@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig. \@ref(community-graph16)).
The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.

@@

![
Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 11 loci. 
Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region. 
Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
](../../results/figures/publication/Figure3Z.pdf){#community-graph width=100%}

@@

The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between white mold screening nurseries and non-white mold screening nurseries after accounting for variation between regions (Table \@ref(tab:amova)).
Because we observed that the sample sizes of wmn isolates and producer isolates were lopsided in several populations (Table \@ref(tab:isolate-table)), we ran the analysis again with regions that had at least ten isolates from both wmn and producer fields, showing similar results (Table \@ref(tab:amova)). 
In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Data).

@@

Table: (\#tab:amova) Analysis of Molecular Variance on Bruvo's genetic distance showing little differentiation between white mold screening nurseries and other sources. The hierarchy was constructed as Source/Region where source is defined as belonging to a white mold screening nursery or not. Bold $\Phi$ values indicate significant values (P < 1.00e^-4^). Numbers in parentheses represent AMOVA run with the regions NE, WA, and MI. S.S. = Sum of Squares, d.f. = degrees of freedom.

@@

Discriminant Analysis of Principal Components (DAPC) was performed grouping by Region with the first 21 principal components, representing 88% of the total variance. 
The first discriminant axis (representing 64% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig. \@ref(DAPC)b). 
The second discriminant axis separated the centroid for the Californian isolates.
The mean population assignment probabilities for all populations with n > 10 showed that only isolates from Mexico, CA, and France had > 50% probabilities of being reassigned to their source populations (Fig. \@ref(DAPC)a). 
DAPC grouping by cultivar used the first 20 principal components, representing 89% of the total variance.
The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were 33.97% (G122), 35.95% (Beryl), and 30.08% (Bunsi).

@@

![
Discriminant Analysis of Principal Components on regions showing that Mexico is differentiated from other populations. 
**(a)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
Numbers next to source populations indicate population size.
All values sum to one.
**(b)** Scatter plot of first two components from DAPC.
Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
The center of each component is represented as grid lines.
](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}

@@

In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a regional level. 
We found significant population differentiation by geographic region, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico. 
To date, with 366 isolates, this is the largest population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
These findings indicate that the white mold screening nurseries can be effective at screening for potential resistant lines within growing regions.

@@

One of the main limitations of this study is the focus on *P. vulgaris* as a host. 
It has been shown that *S. sclerotiorum* in the midwestern United States does not have a particular preference for host [@aldrich-wolfe2015genetic]. 
If the distribution of *S. sclerotiorum* is even across agricultural hosts in the US, then our sample may yet be representative of the genetic pool present in other crops and weedy species.

@@

![
Graph showing complex associations between Mycelial Compatibility Groups (dotted nodes) and Multilocus Genotypes (full nodes). 
Node size reflect the number of samples represented by each node (circle). 
Edges (arrows) point from Mycelial Compatibility Group to Multilocus Genotype where the weight (thickness) of the edge represents the number of samples shared. 
Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing]. 
](../../results/figures/publication/full-graph.pdf){#fullgraph width=100%}

@@

![
Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 16 loci. 
Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region. 
Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
](../../results/figures/publication/FigureS1Z.pdf){#community-graph16 width=100%}

>>>>

<!-- SEE: Avoid sentences that start with the authors/reference. -->
@attanayake2013sclerotinia studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA.
There were no shared microsatellite multilocus genotypes (MLG) between the two populations.
Sexual outcrossing was detected in both populations with higher recombination rate in Chinese population than the USA population.
Additionally, phenotypic traits such as production of sclerotia, mycelial compatibility group (MCG), growth rate were also significantly different between the two populations.
Genetic diversity of *S. sclerotiorum* from soy bean, canola, sunflower, and dry bean in the North Central region of the USA was studied by [@aldrich-wolfe2015genetic].
Microsatellite analysis revealed little host- specificity or genetic-isolation of *S. sclerotiorum* isolates among host crops as 93% of the observed genetic variation was explained by within-host differences.
Isolates from all host crops were primarily reproducing clonally while there were also evidence of genetic recombination of isolates from canola and sunflower.

<!-- Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe. 
In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
No white mold resistant dry bean cultivars exist.
Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic]. -->
To date, no commercial white mold-resistant dry bean cultivars exist. 
Field trials in some regions of the USA have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation; @steadman2004identification].
<!-- This puny paragraph needs a lot of work to more fully describe the operation of the wmn -->

@@

## Data processing and Analysis{-}

All analyses were performed on this data set unless otherwise mentioned.
All data processing and analyses were performed in R version 3.4.1 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif. 
<!-- add something about repeat length consistency  -->
Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of region/source/host/year---meaning that one representative of each genotype was retained for each combined regions, source, host, and year---for subsequent analysis. 
We assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis]. 
To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
Both genotypic diversity and linkage disequilibrium were calculated in *poppr* version 2.4.1 [@kamvar2014poppr].

We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to test for differentiation associated with wmn by quantifying the differentiation conferred by Source and Region [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
Source in this case was defined as a population belonging to a wmn or not.

To identify specific Regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet]. 
This method decomposes the genetic data into principal components, and then uses a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant]. 
To avoid over-fitting data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
This function implements a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximizes the rate of successful group reassignment. 
To assess if cultivar had an influence on genetic diversity between wmn, we first subset the clone-corrected data to contain only samples from wmn and from the cultivars Beryl, Bunsi, and G122 and tested differentiation using AMOVA and DAPC as described above.

### Analysis of Shared Multilocus Genotypes {-}

We wanted to evaluate patterns of connectivity between shared multilocus genotypes across geographic regions.
We first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
Each node was weighted by the fraction of shared MLGs.
Each edge represented a single MLG, but because a single MLG could be present in more than one population, that MLG would be represented by (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
The graph was plotted using the R packages *ggplot2* version 2.2.1 [@ggplot2] and *ggraph* version 1.0.0 [@ggraph].
To ensure that we captured the same community signal, we additionally performed this analysis including the five polymorphic markers described above.

A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in Australia, Mexico, France, and 11 states in the USA (Table \@ref(tab:isolate-table)).
Over 50% of the isolates came from four states, MI (62), ND (60), WA (59), NE (47). 
Four regions had fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1).
We observed 87 MCGs, the most abundant of which ('MCG 5') was represented by 73 isolates over 37 MLGs (Fig \@ref(biggraph)A,C).
After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with 318 isolates representing 165 MLGs (Table \@ref(tab:iatab)).
Analysis of the genotype accumulation curve showed no clear plateau (Supplemental Information), indicating that we would likely obtain more multilocus genotypes if we were to genotype more loci. The number of observed alleles per locus ranged from two to 10 with an average of 6.27 (Table \@ref(tab:locus-stats)).

@@

@@

The forward-backward selection process of the dbRDA models on clone-corrected data revealed Year, Region, Host, and Aggressiveness to be the optimal variables for the reduced model, accounting for 45% of the total variation.
ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.0674 (*P* = 0.001).
Assessment of the marginal effects showed that region, host, year, and aggressiveness significantly explained genetic variation (*P* = 0.001, *P* = 0.009 for region).
It was revealed that there was multicollinearity when MCG was combined with any other variable, so we followed the same procedure, dropping MCG from the list of potential predictors. 
From these results, we found that aggressiveness, year, region, and host were optimal, accounting for 17.6% of the total variation.
ANOVA revealed significant effects with an adjusted R^2^ of 0.0325 (*P* = 0.001).
While the marginal effect assessment showed that region, host, and year all significantly explained the variation at *P* = 0.001, and aggressiveness significantly explained the variation at *P* = 0.045.
Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
Variance partitioning of the independent variables without MCG indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.

@@

![
Biplot showing five most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
Open circles represent the 318 clone-corrected samples in ordination space. 
](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=61%}

@@

![
Boxplot representing aggressiveness (vertical axis) by population (horizontal axis) arranged in descending order of mean aggressiveness for all populations with N > 5. Box boundaries represent 25% and 75% percentiles and whiskers extend to at most 1.5 $\times$ the interquartile range. Horizontal line represents median value. Circles represent individual observations where filled circles indicate isolates from white mold screening nurseries (wmn).
](../../results/figures/publication/aggressiveness.pdf){#aggressiveness-boxplot}

@@

Our assessment of aggressiveness in association with Region showed a significant effect (P < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig. \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)). 
MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 (MCG 44) to 4.6 (MCG 49).
We additionally found a significant effect for MLGs (*P* = 7.44e^-4^) for MLGs, with means that ranged from 6.0 (MLG 78) to 4.3 (MLG 140).

@@

Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set (Fig. \@ref(fullgraph)).
The three communities with the most members contained eight of the 10 most abundant MCGs. 
Comparing these communities with Bruvo's genetic distance revealed an average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
When we assessed the number of times two different MLGs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of 16.5 potential heterothallic parings per MLG. 
Representing just four isolates, 'MLG 75' had 57 neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
Overall, there was no clear pattern to the association between MLG and MCGs.

@@

![
Data show complex associations between Mycelial Compatibility Groups and Multilocus Genotypes.
**A)** Barplot of Mycelial Compatibility Group (MCG) abundance in descending order.
Singletons (46) were truncated, leaving 41 MCGs.
White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
The transparency of the bars represent the evenness of the distribution of the MLGs within a given MCG.
A dashed box surrounds the 8 most common MCGs representing >51% of the data.
**B)** Full graph-representation of the relationship between MCGs (open circles) and MLGs (filled circles).
Details in fig. \@ref(fullgraph). 
**C)** A subset of **B** representing the 8 most common MCGs and their associated MLGs (dashed box in **A**).
Filled nodes represent MLGs and open nodes represent MCGs.
Numbers inside nodes are the MLG/MCG label. 
Node (circle) area scaled to the number of samples represented (range: 1--73).
Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of shared isolates (range: 1--19).
Edges extending from MLGs displayed to other MCGs are not shown.
](../../results/figures/publication/mcg-subgraph-with-context.pdf){#biggraph width=100%}

@@

## Structure of Shared Multilocus Genotypes {-}

@@

We observed a total of 165 MLGs (215 with 16 loci).
The most abundant MLG was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five regions (NE, MI, WA, CO, and ND). 
Within regions, the genotypes were relatively evenly distributed with moderate to high diversity (Table \@ref(tab:iatab)). 
Of the 165 MLGs, 76 (46%) were found in at least two geographic regions except those found in WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig. \@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a midwest community (CO, NE, ND, MI), and an international community (MN, France, Australia).
Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig. \@ref(community-graph16)).
The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.

@@

![
Network of populations (nodes/circles) and their shared multilocus genotypes (MLG) (edges/lines) genotyped over 11 loci. 
Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region (bottom legend). 
Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
](../../results/figures/publication/mlg-11.pdf){#community-graph width=100%}

@@

The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between wmn and non-wmn after accounting for variation between regions (Table \@ref(tab:amova)).
Because we observed that the sample sizes of wmn isolates and producer isolates were lopsided in several populations (Table \@ref(tab:isolate-table)), we ran the analysis again with regions that had at least ten isolates from both wmn and producer fields, showing similar results (Table \@ref(tab:amova)). 
In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Data).

@@

Table: (\#tab:amova) Analysis of Molecular Variance on Bruvo's genetic distance showing little differentiation between white mold screening nurseries (wmn) and other sources. The hierarchy was constructed as Source/Region where source is defined as belonging to a wmn or not. Bold $\Phi$ values indicate significant values (P < 1.00e^-4^). Numbers in parentheses represent AMOVA run with the regions NE, WA, and MI. S.S. = Sum of Squares, d.f. = degrees of freedom.

@@

Discriminant Analysis of Principal Components (DAPC) was performed grouping by Region with the first 21 principal components, representing 88% of the total variance. 
The first discriminant axis (representing 64% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig. \@ref(DAPC)b). 
The second discriminant axis separated the centroid for the California isolates.
The mean population assignment probabilities for all populations with n > 10 showed that only isolates from Mexico, CA, and France had > 50% probabilities of being reassigned to their source populations (Fig. \@ref(DAPC)a). 
DAPC grouping by cultivar used the first 20 principal components, representing 89% of the total variance.
The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were 34.0% (G122), 36.0% (Beryl), and 30.1% (Bunsi).

@@

![
Discriminant Analysis of Principal Components on regions showing that Mexico is differentiated from other populations. 
**A)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
Numbers next to source populations indicate population size.
All values sum to one.
**B)** Scatter plot of first two components from DAPC.
Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
The center of each component is represented as grid lines.
](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}

@@

In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a within regions. 
We found significant population differentiation by geographic region, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico. 
To date, with 366 isolates, this is the largest single population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
These findings indicate that the wmns can be effective at screening for potential resistant lines within growing regions.

@@

One of the main limitations of this study is the focus on *P. vulgaris* as a host. 
It has been shown that *S. sclerotiorum* in the midwestern United States does not have a particular preference for host [@aldrich-wolfe2015genetic]. 
If the distribution of *S. sclerotiorum* is even across agricultural hosts in the USA, then our sample may yet be representative of the genetic pool present in other crops and weedy species.

@@

![
Graph showing complex associations between Mycelial Compatibility Groups (MCG) (dotted nodes) and Multilocus Genotypes (MLG) (full nodes) where the number in each node represents the MLG/MCG assignment. 
Node size reflect the number of samples represented by each node (circle). 
Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of samples shared. 
Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing]. 
](../../results/figures/publication/full-graph.pdf){#fullgraph width=100%}

@@

![
Network of populations (nodes/circles) and their shared multilocus genotypes (MLG) (edges/lines) genotyped over 16 loci. 
Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region (bottom legend). 
Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
](../../results/figures/publication/mlg-16.pdf){#community-graph16 width=100%}
